This PDF is a Git Repository\\Containing its Own \LaTeX\ Source\\and a Copy of Itself

### 2017-05-03 article.tex
I wonder how many readers will think I'm British?
Minor wording tweaks

The file format for Git bundles doesn't appear to be formally
specified anywhere, however, inspecting \texttt{bundle.c} reveals that
it's relatively straightforward:
\def\returnkey{\hspace*{1pt}{\setbox9=\hbox{$\hookleftarrow$}\tikz\node[text width=\wd9,fill=lightgray,rounded corners=1pt] at (0,0){$\hookleftarrow$};}\hspace*{1pt}}
\begin{center}
\begin{tikzpicture}
  \node[text width=0.9\hsize] at (0,0) (sig) {\footnotesize\tt\# v2 git bundle{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.9\hsize] (sigbox) {};
  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\footnotesize Git Bundle Signature};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (sig.south west) (b1) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b1.south west) (b2) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b2.south west) (b3) {\footnotesize\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b3.south west) (b4) {\footnotesize\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.9\hsize] (digestbox) {};
  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\footnotesize Digest};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b4.south west) (empty) {{\tiny\returnkey}\footnotesize\ };
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (empty.south west) (pack) {{\footnotesize\tt PACK}$\ldots$};
  \node[inner sep=0mm,fit=(pack),draw,rounded corners,dashed,minimum width=0.9\hsize] (packbox) {};
  \node[draw,rounded corners,fill=white] at (pack.south) {\footnotesize Git Packfile};
\end{tikzpicture}
\end{center}
Git uses its Packfile format to compress the objects in its database,
as well as to reduce network bandwidth when pushing and pulling, which
makes it an obvious choice for use inside bundles. This of course
raises the question: What is the format for a Git Packfile?

@@

Okay, the next step is to figure out the ``chunk'' format.  The chunk
header is variable length, and can be as small as one byte. It encodes
the object's type and its \emph{uncompressed} size. If the object is
a \textit{delta} (\textit{i.e.}, a diff, as opposed to a complete
object), the header is followed by either the SHA-1 hash of the base
object to which the delta should be applied, or a byte reference
within the packfile for the start of the base object. The remainder of
the chunk consists of the object data, zlib-compressed.

@@

Where can we put a 5-byte DEFLATE header such that it won't corrupt
the PDF? We could use our standard trick of putting it in a PDF object
stream that we've exploited countless times before to enable
PoC$\|$GTFO polyglots. The trouble with that is that object streams
are fixed-length, so once the PDF is decompressed (\textit{i.e.},~when
a repo is cloned from the git bundle), then all of the 5-byte DEFLATE
headers will disappear and the object stream lengths would all be
incorrect. Instead, I chose to use PDF comments, which start at any
occurrence of the percent sign character~(\%) outside a string or
stream and continue until the first occurrence of a newline.  All of
the PDF viewers I tested don't seem to care if comments include
non-ASCII characters; they seem to simply scan for a
newline. Therefore, we can inject ``\texttt{\%\textbackslash n}''
between PDF objects and move the DEFLATE headers there. The only
caveat is that the DEFLATE header itself can't contain a newline
byte~(\texttt{0x0A}), otherwise the comment would be ended
prematurely.  We can resolve that by adding extra spaces to the end of
the comment, increasing the length of the following DEFLATE block and
thus increasing the length bytes in the DEFLATE header and avoiding
the \texttt{0x0A}. The only concession made with this approach is that
PDF Xref offsets in the deflated version of the PDF will be off by a
multiple of 5, due to the removed DEFLATE headers. Fortunately, most
PDF readers can gracefully handle incorrect Xref offsets (at the
expense of a slower loading time), and this will only affect the PDF
contained in the repository, \emph{not} the PDF polyglot.

@@

At this point, we have all the tools necessary to create a generalized
PDF/Git Bundle polyglot for \emph{almost} any PDF and git
repository. The only remaining hurdle is that some viewers require
that the PDF occur as early in the packfile as possible. At first, I
considered applying another patch directly the git source code to make
the uncompressed object first in the packfile. This approach proved to
be very involved, in part due to git's UNIX design philosophy and
architecture of generic code reuse. We're already updating the
packfile's SHA-1 hash due to changing the DEFLATE headers, so instead
I decided to simply reorder the objects after-the-fact, subsequent to
the DEFLATE header fix but before we update the hash. The only
challenge is that moving objects in the packfile has the potential to
break offset delta objects, since they refer to their base objects via
a byte offset within the packfile.  Moving the PDF to the beginning
will break any offset delta objects that occur after the original
position of the PDF which refer to base objects that occur before the
original position of the PDF. That's why I had to dive deeper into the
rabbit hole of the packfile format to understand the delta object
headers (as you saw at the end of Section~\ref{sec:GitBundleFormat},
if you were brave enough to finish it), since I originally attempted
to rewrite the broken ones. Rewriting the broken offset delta objects
is the \emph{correct} solution, but, in the end, I discovered a much
simpler way.

>>>>

The file format for Git bundles doesn't appear to be formally
specified anywhere, however, inspecting \texttt{bundle.c} reveals that
it's relatively straightforward:
\def\returnkey{\hspace*{1pt}{\setbox9=\hbox{$\hookleftarrow$}\tikz\node[text width=\wd9,fill=lightgray,rounded corners=1pt] at (0,0){$\hookleftarrow$};}\hspace*{1pt}}
\begin{center}
\begin{tikzpicture}
  \node[text width=0.9\hsize] at (0,0) (sig) {\footnotesize\tt\# v2 git bundle{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.9\hsize] (sigbox) {};
  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\footnotesize Git Bundle Signature};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (sig.south west) (b1) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b1.south west) (b2) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b2.south west) (b3) {\footnotesize\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b3.south west) (b4) {\footnotesize\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.9\hsize] (digestbox) {};
  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\footnotesize Digest};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b4.south west) (empty) {{\tiny\returnkey}\footnotesize\ };
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (empty.south west) (pack) {{\footnotesize\tt PACK}$\ldots$};
  \node[inner sep=0mm,fit=(pack),draw,rounded corners,dashed,minimum width=0.9\hsize] (packbox) {};
  \node[draw,rounded corners,fill=white] at (pack.south) {\footnotesize Git Packfile};
\end{tikzpicture}
\end{center}
Git has another custom format called a \emph{Packfile} that it uses to
compress the objects in its database, as well as to reduce network
bandwidth when pushing and pulling. The packfile is therefore an
obvious choice for storing objects inside bundles. This of course
raises the question: What is the format for a Git Packfile?

@@

Right, the next step is to figure out the ``chunk'' format.  The chunk
header is variable length, and can be as small as one byte. It encodes
the object's type and its \emph{uncompressed} size. If the object is
a \textit{delta} (\textit{i.e.}, a diff, as opposed to a complete
object), the header is followed by either the SHA-1 hash of the base
object to which the delta should be applied, or a byte reference
within the packfile for the start of the base object. The remainder of
the chunk consists of the object data, zlib-compressed.

@@

Where can we put a 5-byte DEFLATE header such that it won't corrupt
the PDF? We could use our standard trick of putting it in a PDF object
stream that we've exploited countless times before to enable
PoC$\|$GTFO polyglots. The trouble with that is: Object streams are
fixed-length, so once the PDF is decompressed (\textit{i.e.},~when a
repo is cloned from the git bundle), then all of the 5-byte DEFLATE
headers will disappear and the object stream lengths would all be
incorrect. Instead, I chose to use PDF comments, which start at any
occurrence of the percent sign character~(\%) outside a string or
stream and continue until the first occurrence of a newline.  All of
the PDF viewers I tested don't seem to care if comments include
non-ASCII characters; they seem to simply scan for a
newline. Therefore, we can inject ``\texttt{\%\textbackslash n}''
between PDF objects and move the DEFLATE headers there. The only
caveat is that the DEFLATE header itself can't contain a newline
byte~(\texttt{0x0A}), otherwise the comment would be ended
prematurely.  We can resolve that by adding extra spaces to the end of
the comment, increasing the length of the following DEFLATE block and
thus increasing the length bytes in the DEFLATE header and avoiding
the \texttt{0x0A}. The only concession made with this approach is that
PDF Xref offsets in the deflated version of the PDF will be off by a
multiple of 5, due to the removed DEFLATE headers. Fortunately, most
PDF readers can gracefully handle incorrect Xref offsets (at the
expense of a slower loading time), and this will only affect the PDF
contained in the repository, \emph{not} the PDF polyglot.

@@

At this point, we have all the tools necessary to create a generalized
PDF/Git Bundle polyglot for \emph{almost} any PDF and git
repository. The only remaining hurdle is that some viewers require
that the PDF occur as early in the packfile as possible. At first, I
considered applying another patch directly the git source code to make
the uncompressed object first in the packfile. This approach proved to
be very involved, in part due to git's UNIX design philosophy and
architecture of generic code reuse. We're already updating the
packfile's SHA-1 hash due to changing the DEFLATE headers, so instead
I decided to simply reorder the objects after-the-fact, subsequent to
the DEFLATE header fix but before we update the hash. The only
challenge is that moving objects in the packfile has the potential to
break offset delta objects, since they refer to their base objects via
a byte offset within the packfile.  Moving the PDF to the beginning
will break any offset delta objects that occur after the original
position of the PDF that refer to base objects that occur before the
original position of the PDF. That's why I had to dive deeper into the
rabbit hole of the packfile format to understand the delta object
headers (as you saw at the end of Section~\ref{sec:GitBundleFormat},
if you were brave enough to finish it), since I originally attempted
to rewrite the broken ones. Rewriting the broken offset delta objects
is the \emph{correct} solution, but, in the end, I discovered a much
simpler way.
